Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-09 manuscript_210505.tex
Deleted most old figures and large excess files

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{CJK}
\usepackage{framed}
%\usepackage[hidelinks]{hyperref}
%\usepackage{algorithm}
%\usepackage{algpseudocode}
%\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}

@@

\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\rood}[1]{#1}
\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
%\newcommand*{\blauw}[1]{#1}
\newcommand*{\groen}[1]{\textcolor{green}{#1}}
%\newcommand*{\groen}[1]{#1}
%\newcommand*{\blauwr}[1]{{\color{blue}{#1}}}
%\newcommand*{\blauwr}[1]{#1}

\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}

\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes on
%\newcommand*{\noter}[1]{}					% notes off

@@

\textbf{\Large NOTES}
\begin{itemize}
\item All figs as pdf (makes compilation faster) or tiff
\item For quotation marks in tex use ``XXX'' rather than "XXX"
\item Strong structure is critical

-- It must be a narrative story with a logical flow -- reader must understand where you are taking them next and why

-- Topic sentences

-- Everything is a funnel

\item Figures are included directly after the paragraph after which they are first mentioned

\item We use hybridization / association and dehybridization / dissociation interchangeably. Maybe this is OK, or maybe we should pick one terminology and stick with it.

\item In first mention of anything in SI say, e.g., Fig.~S1 in the Supporting Information. From then on, can just say Fig.~SX.

\item Write out in words numbers less than 10, e.g., one-base, two-times, seven, and use numerals for 10 and greater, e.g. 10-mer, 15-fold, 42 (Only exception I can think of is reporting ranges of numbers, e.g., 2-6.)

\item Every figure caption also starts with a topic sentence. Should also contain, in general, a takeaway message for the reader in addition to a description of the figure content.

\item We probably also need a table of contents (TOC) image. This should be emblematic of the paper as a whole, and cannot be just (any portion of) an existing image. Insert this image at the end of the article after the references.

\end{itemize}

@@

\noindent A robust understanding of the sequence-dependent thermodynamics of DNA hybridization has enabled rapid advances in DNA nanotechnology. A fundamental understanding of the sequence-dependent kinetics and mechanisms of hybridization and dehybridization remains comparatively underdeveloped. In this work, we establish new understanding of the sequence-dependent hybridization/dehybridization kinetics and mechanism within a family of self-complementary pairs of 10-mer DNA oligomers by integrating coarse-grained molecular simulation, machine learning of the slow dynamical pathways, data-driven inference of long-time kinetic models, and experimental temperature-jump infrared spectroscopy. For a repetitive ATATATATAT sequence, we resolve a rugged dynamical landscape comprising multiple metsatable states, numerous competing hybridization/dehybridization pathways, and a spectrum of dynamical relaxations. Introduction of a G:C pair at the terminus (GATATATATC) or center (ATATGCATAT) of the sequence reduces the ruggedness of the dynamics landscape by eliminating a number of metastable states and reducing the number of competing dynamical pathways. Only by introducing a G:C pair midway between the terminus and the center to maximally disrupt the repetitive nature of the sequence (ATGATATCAT) do we recover a canonical two-state model of hybridization/dehybridization over a smooth landscape with no intermediate metastable states. Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation in burgeoning DNA nanotechnology applications.

@@

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnological applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TM} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the \textit{dynamical}, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation For Imaging In Nanoscale Topography), but developments in dynamic DNA nanotechnologies have outpaced our understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena from the picosecond to millisecond range.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} \rood{and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence.} \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence} \noter{Is this red clause correct? Are there simulation components to the work? Was it on/off rates that were predicted?} \rood{A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved.} \noter{Accurate? More here? Does this adequately convey wrt to both citations why the two cited investigations are not the final word. i.e., why is there still a need for the present paper -- why is the problem not solved by these works, what did they leave unanswered?} For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates} \noter{Should these references also be included here? \citep{Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019}} Out-of-register ``shifted'' base paired structures \citep{Phys2014, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the extent to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics has been built from decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated equilibrium analysis, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Phys2014} \noter{I moved this citation here -- is that OK?} but they can be observed by employing enhanced sampling techniques such as metadynamics, umbrella sampling,  transition path sampling, and forward flux sampling,\citep{Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy. \addref{Girsanov (Keller ? 10.1063/1.4989474, 10.1063/1.5027728 ,10.1137/17M1124772), TRAM (Noe, Hummer ? 10.1073/pnas.1525092113), and MBAR for paths (Chodera, Shirts, Noe ? 10.1063/1.3592152, 10.1063/1.3592153)} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales \noter{micro is OK? milli?} via unbiased coarse-grained molecular dynamics (CGMD) simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to a speedup factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to \rood{do X, Y, and Z}. \addref{}

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family of oligomers -- 5$^\prime$ATATATATAT3$^\prime$ (AT-all), 5$^\prime$GATATATATC3$^\prime$ (GC-end), 5$^\prime$ATATGCATAT3$^\prime$ (GC-core), and 5$^\prime$ATGATATCAT3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of a G:C pair within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new and existing experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Phys2014,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of non-native states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states that facilitate the hybridization/dehybridization process. Analysis of the configurational ensemble within each metastable state reveals the conformational heterogeneity and dynamical interconversions proceeding on time scales that lie below those resolvable within the long-time kinetic model. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to sculpt and qualitatively alter the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT and DNA barcoding. \addref{} \noter{Is barcoding appropriate here, or just PAINT is sufficient?}

@@

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by ultra-fast temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$ATATATATAT3$^\prime$ (AT-all), 5$^\prime$GATATATATC3$^\prime$ (GC-end), 5$^\prime$ATATGCATAT3$^\prime$ (GC-core), and 5$^\prime$ATGATATCAT3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSummary, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

All calculations were performed using the LAMMPS simulation package in accordance with best practices for the 3SPN.2 model. \addref{LAMMPS} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{XX} mol/L. Solvent effects were modeled implicitly by employing Langevin dynamics \citep{Schneider1978Molecular-dynamicsTransitions} \noter{Is this the right reference for Langevin dynamics?} with an experimentally motivated per-site friction coefficient of $\xi$ = 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Hinckley2013AnHybridization} \addref{A. E. Nkodo, G. M. Garnier, B. Tinland, C. Desruisseaux, L. C. Mc-Cormick, G. Drouin, and G. W. Slater,Electrophoresis22, 2424 (2001)} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \addref{Debye-Huckel} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \addref{Langevin thermostat} Each sequence was simulated at its melting temperature -- AT-all: \rood{XX} K, GC-end: \rood{XX} K, GC-core: \rood{XX} K, GC-mix: \rood{XX} K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by \rood{XXX}. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \addref{G. Bussi and M. Parrinello,Phys.Rev.E75, 056707 (2007).} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. Initial system configurations were prepared \rood{How? How specify initial configurations and apply any energy minimization?} Initial bead velocities assigned from a Maxwell-Boltzmann distribution at the system temperature. \addref{Maxwell-Boltzmann} Each simulation was conducted for 26 $\mu$s and frame saved to disc every 100 ps, requiring $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

%A featurization of trajectories was required to capture fundamental dynamics while accounting for translational and rotational symmetries. Pairwise distances between base centers of mass -- similar to heavy atom distances in protein models \citep{Sidky}-- satisfied these requirements and were calculated at each frame using the MDtraj Python library.\citep{McGibbon2015MDTraj:Trajectories} Although pairwise distances remove translational and rotational symmetries, additional symmetries arise from the self-complementary nature of these sequences -- the sense and anti-sense strands in each pair are identical to each other. To account for this, we averaged permutable distances (45 pairs in total) together into a ``symmetrized'' feature set following a similar procedure used in tICAgg coordinate construction.\citep{Sengupta2019AutomatedSelf-assembly} The VAMP-2 scoring method was employed to evaluate the kinetic variance of the feature set and optimize hyperparameters.\citep{WuVariationalData, Mardt2018VAMPnetsKinetics} Additional details on the VAMP-2 score can be found in the supplemental information. By comparing scores across several candidate feature sets (Figure \ref{fig:allseq_features_vamp2}), we found that minimal kinetic information was lost in the  symmetrization procedure. Furthermore, the smaller feature set enabled faster training times and better statistics over permutable distances, without appearing to suffer a loss in generality or model resolution.

@@

\bibliography{references}
%\bibliography{refs_mendeley2}

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{CJK}
\usepackage{framed}

%add refs to SI
%\usepackage{xr}
%\externaldocument{SI}

%\usepackage[hidelinks]{hyperref}
%\usepackage{algorithm}
%\usepackage{algpseudocode}
%\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}

@@

\newcommand*{\rood}[1]{{\color{red}{#1}}}
\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
\newcommand*{\groen}[1]{\textcolor{green}{#1}}

\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off

@@

@@

\noindent A robust understanding of the sequence-dependent thermodynamics of DNA hybridization has enabled rapid advances in DNA nanotechnology. A fundamental understanding of the sequence-dependent kinetics and mechanisms of hybridization and dehybridization remains comparatively underdeveloped. In this work, we establish new understanding of the sequence-dependent hybridization/dehybridization kinetics and mechanism within a family of self-complementary pairs of 10-mer DNA oligomers by integrating coarse-grained molecular simulation, machine learning of the slow dynamical pathways, data-driven inference of long-time kinetic models, and experimental temperature-jump infrared spectroscopy. For a repetitive ATATATATAT sequence \noteb{not sure if we want to use 5$^\prime$-ATATATATAT-3$^\prime$ notation here as well?}, we resolve a rugged dynamical landscape comprising multiple metsatable states, numerous competing hybridization/dehybridization pathways, and a spectrum of dynamical relaxations. Introduction of a G:C pair at the terminus (GATATATATC) or center (ATATGCATAT) of the sequence reduces the ruggedness of the dynamics landscape by eliminating a number of metastable states and reducing the number of competing dynamical pathways. Only by introducing a G:C pair midway between the terminus and the center to maximally disrupt the repetitive nature of the sequence (ATGATATCAT) do we recover a canonical two-state model of hybridization/dehybridization over a smooth landscape with no intermediate metastable states. Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation in burgeoning DNA nanotechnology applications.

@@

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnological applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TM} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the \textit{dynamical}, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation For Imaging In Nanoscale Topography), but developments in dynamic DNA nanotechnologies have outpaced our understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena from the picosecond to millisecond range.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence. \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence, Shah2019}. A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved. For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019} Out-of-register ``shifted'' base paired structures \citep{Phys2014, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the extent to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics has been built from decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated equilibrium analysis, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Phys2014} but they can be observed by employing enhanced sampling techniques such as metadynamics, umbrella sampling,  transition path sampling, and forward flux sampling,\citep{Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} \noteb{these references correspond to CG enhanced sampling studies, I can refer to enhanced sampling on AA here instead (\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations}) but I think the CG studies should be mentioned since not much has been done for AA hybridization} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy.~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics (CGMD) simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to a speedup factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacements \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami structure \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}.

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family of oligomers -- 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and 5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of a G:C pair within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new and existing experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Phys2014,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of non-native states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states that facilitate the hybridization/dehybridization process. Analysis of the configurational ensemble within each metastable state reveals the conformational heterogeneity and dynamical interconversions proceeding on time scales that lie below those resolvable within the long-time kinetic model. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to sculpt and qualitatively alter the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}

@@

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by ultra-fast temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC3-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSummary, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997Short-RangeDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of 7 mol/L. Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers.} with an experimentally motivated per-site friction coefficient of $\xi$ = 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. Each sequence was simulated at its melting temperature -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the Nearest Neighbors model predicted melting temperature. The temperature at which the hybridized and dissociated populations were closest was selected. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was set, according to 3SPN.2 documentation, to the coordinates of Arnott et al.\ ~\citep{Arnott1976CRCBiology} The dissociated state was generated by shifting one oligomer from the hybridized state by 1 nm in the x, y, and z directions. The magnitude of this shift did not affect the results so long as all native WC bonds were sufficiently separated. \noteb{No energy minimization (no 3spn2 precedent on this). The lammps default emin techniques may not work for this type of CG forcefield. Given the sampling it seems that  1 $\mu$s of equilibration should be sufficient?} Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the system temperature. \addref{Maxwell-Boltzmann --- \noteb{not sure what to use here, usually not cited?}} Each simulation was conducted for 26 $\mu$s and frame saved to disc every 100 ps, requiring $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

%A featurization of trajectories was required to capture fundamental dynamics while accounting for translational and rotational symmetries. Pairwise distances between base centers of mass -- similar to heavy atom distances in protein models \citep{Sidky}-- satisfied these requirements and were calculated at each frame using the MDtraj Python library.\citep{McGibbon2015MDTraj:Trajectories} Although pairwise distances remove translational and rotational symmetries, additional symmetries arise from the self-complementary nature of these sequences -- the sense and anti-sense strands in each pair are identical to each other. To account for this, we averaged permutable distances (45 pairs in total) together into a ``symmetrized'' feature set following a similar procedure used in tICAgg coordinate construction.\citep{Sengupta2019AutomatedSelf-assembly} The VAMP-2 scoring method was employed to evaluate the kinetic variance of the feature set and optimize hyperparameters.\cite{ Mardt2018VAMPnetsKinetics} Additional details on the VAMP-2 score can be found in the supplemental information. By comparing scores across several candidate feature sets (Figure \ref{fig:allseq_features_vamp2}), we found that minimal kinetic information was lost in the  symmetrization procedure. Furthermore, the smaller feature set enabled faster training times and better statistics over permutable distances, without appearing to suffer a loss in generality or model resolution.

@@

%\bibliographystyle{unsrt}
\bibliography{references_210505}

%\bibliography{refs_mendeley2}
